### 2021-12-24 chapters/9_discussion.tex
background
draft paper 0.1

\chapter{Discussion}
\label{ch:discussion}
In this chapter, we discuss the results of our experiments on clustering mutants.
First, we discuss the results of the white box approach. 
The results are discussed as one and are compared between the variations(with Levenshtein distance characteristic and without).
Second we discuss the results of the black box approach and compare them to those of the white box approach.
And last we discuss the practicality of our research.

@@

This includes both variations of the experiment.

@@

In both variations of the experiment we observe a reduction in average accuracy when increasing the performance by reducing the number of clusters.

@@

\section{RQ 2: Training a machine learning model and clustering mutant characteristics}
The maximum deviation from the original mutation score, for all variants(with and without Levenshtein distance characteristic) of the experiment, is above the statistical threshold.
The biggest deviation measured is 22.39\%.
This is significantly higher than the biggest deviation measured in the first experiment.
56\% of the results of both variants had a deviation in mutation score bigger than 5\%.
\begin{finding}
Our hypothesis does not hold for the clustering the mutants with the \acrshort{fcm} model trained in our experiment.
\end{finding}
We observe that the number of clusters does not have an obvious relation to the total amount of mutants. 
This is expected as the subtractive clustering does not look at the total amount of data points(mutants) but to the contents(characteristics) of the data points.
The number of clusters compared to the that of the hierarchical clustering is significantly lower.
While we do not know the exact clusters centroids the subtractive model calculated, a low number of clusters indicates close relationships between certain mutants.
\begin{finding}
The mutants represented as characteristics in our data set per project have, according to the subtractive clustering model, close relationships to certain other mutants in the same data set.
\end{finding}
We observe that the minimum lowest minimum cluster accuracy measured over all variants and projects is 50.54\% with and average of 62\%. 
This is higher than the minimum of the hierarchical clustering.
This means that in all project there was not one cluster that was evenly divided.
While the minimum accuracy was higher, the maximum accuracy was lower compared to the maximum accuracy of the hierarchical clustering results.
The highest accuracy measured over all variants and projects is 97.80\% with an average of 83.79\%
This means that in all project there was not one cluster that consisted of only survived or killed mutants.
\begin{finding}
The overall accuracy of the clusters generated by the \acrshort{fcm} model in both variants is lower compared to that of the hierarchical clustering.
\end{finding}
When we compare the accuracy per project between the \acrshort{fcm} model and hierarchical clustering we can observe that the lower accuracy of the \acrshort{fcm} model clusters have a higher impact on the deviation in mutation score.
This is caused by the relatively low number of clusters.
One cluster contains more mutants which results in a bigger weight per cluster.
\begin{finding}
The inaccuracy of clusters has a bigger impact on mutation score when the number of clusters is lower.
\end{finding}

\section{Application}
The application of our research is limited to all software projects written in the language Java and the requirements of PIT\cite{pit}.
We can repeat our experiment for every java project that can be run by PIT.
The scope of the research done in the domain of mutation clustering(Chapter \ref{ch:related_work}) is a subset of the scope in our research.
With our research it is possible to cluster all the mutants used in other related research.
The extraction of mutants is build within the PIT plugin system which is a plug and play system.
Applying our research can be done in three steps.
The first step is to run PIT with the plugin to gather the characteristics.
Gathering the characteristics does not take more than a few minutes depending on the number of mutants generated.
The plugin also makes sure PIT exits without executing mutants after gathering the characteristics.
The second step is to run the script for either the \acrshort{fcm} model or the hierarchical clustering.
The last step is to run PIT again with the plugin to filter the mutants based on the results of the clustering.
It is possible to build all these steps into a DevOps pipeline to automate the process. 
The plugins and scripts needed for these steps are available in the code repository on Github\cite{rbasarat-repo}.

\section{Threats to validity}
In this section, we discuss the threats to internal validity, external validity and reliability.
\subsection{Internal validity}
\subsubsection{Equivalent mutants}
Equivalent mutants are mutants that are syntactically different but functionally identical.
These mutants influence the mutation score while giving the same result as their equivalent counterpart.
In our research equivalent mutants could give a false image of the accuracy of the clusters and the overall accuracy if such a mutant was randomly selected.
PIT contains some functionality for reducing equivalent mutants\cite{pit}.
We also reduced the chance of inaccuracy by repeating our experiment.

>>>>

\chapter{Discussion}
\label{ch:discussion}
In this chapter, we discuss the results of our experiments on clustering mutants.
First, we discuss the results of the white box approach. 
The results are discussed as one and are compared between the variants(with Levenshtein distance characteristic and without).
Second we discuss the results of the black box approach and compare them to those of the white box approach.
And last we discuss the practicality of our research.

@@

This includes both variants of the experiment.

@@

In both variants of the experiment we observe a reduction in average accuracy when increasing the performance by reducing the number of clusters.

@@

\section{RQ 2: Training a machine learning model and clustering mutant characteristics}
The maximum deviation from the original mutation score, for all variants(with and without Levenshtein distance characteristic) of the experiment, is above the statistical threshold.
We can observe that the distance between whiskers of the box-plots is more than the \textit{alpha-value} of 0.05.
Thus we cannot reject the null hypothesis and accept our alternative hypothesis.
The biggest deviation measured is 22.39\%.
This is significantly higher than the biggest deviation measured in the first experiment.
56\% of the results of both variants had a deviation in mutation score bigger than 5\%.
\begin{finding}
Our hypothesis does not hold for the clustering the mutants with the \acrshort{fcm} model trained in our experiment.
\end{finding}
We observe that the number of clusters does not have an obvious relation to the total amount of mutants. 
This is expected as the subtractive clustering does not look at the total amount of data points(mutants) but to the contents(characteristics) of the data points.
The number of clusters compared to the that of the hierarchical clustering is significantly lower.
While we do not know the exact clusters centroids the subtractive model calculated, a low number of clusters indicates close relationships between certain mutants.
\begin{finding}
The mutants represented as characteristics in our data set per project have, according to the subtractive clustering model, close relationships to certain other mutants in the same data set.
\end{finding}
We observe that the minimum lowest minimum cluster accuracy measured over all variants and projects is 50.54\% with and average of 62\%. 
This is higher than the minimum of the hierarchical clustering.
This means that in all project there was not one cluster that was evenly divided.
While the minimum accuracy was higher, the maximum accuracy was lower compared to the maximum accuracy of the hierarchical clustering results.
The highest accuracy measured over all variants and projects is 97.80\% with an average of 83.79\%
This means that in all project there was not one cluster that consisted of only survived or killed mutants.
\begin{finding}
The overall accuracy of the clusters generated by the \acrshort{fcm} model in both variants is lower compared to that of the hierarchical clustering.
\end{finding}
When we compare the accuracy per project between the \acrshort{fcm} model and hierarchical clustering we can observe that the lower accuracy of the \acrshort{fcm} model clusters have a higher impact on the deviation in mutation score.
This is caused by the relatively low number of clusters.
One cluster contains more mutants which results in a bigger weight per cluster.
\begin{finding}
The inaccuracy of clusters has a bigger impact on mutation score when the number of clusters is lower.
\end{finding}

\section{Application}
The application of our research is limited to all software projects written in the language Java and the requirements of PIT\cite{pit}.
We can repeat our experiment for every java project that can be run by PIT.
The scope of the research done in the domain of mutation clustering(Chapter \ref{ch:related_work}) is a subset of the scope in our research.
With our research it is possible to cluster all the mutants used in other related research.
The extraction of mutants is build within the PIT plugin system which is a plug and play system.
Applying our research can be done in three steps.
The first step is to run PIT with the plugin to gather the characteristics.
Gathering the characteristics does not take more than a few minutes depending on the number of mutants generated.
The plugin also makes sure PIT exits without executing mutants after gathering the characteristics.
The second step is to run the script for either the \acrshort{fcm} model or the hierarchical clustering.
The last step is to run PIT again with the plugin to filter the mutants based on the results of the clustering.
It is possible to build all these steps into a DevOps pipeline to automate the process. 
The plugins and scripts needed for these steps are available in the code repository on Github\cite{rbasarat-repo}.
The plugins can be found in the directory \textit{pitest-clustering-plugin}.
The scripts used can be found in the directories \textit{FCM-clustering} and \textit{hierarchical-clustering}.

\section{Threats to validity}
In this section, we discuss the threats to internal validity and external validity.
\subsection{Internal validity}
\subsubsection{Equivalent mutants}
Equivalent mutants are mutants that are syntactically different but functionally identical.
These mutants influence the mutation score while giving the same result as their equivalent counterpart.
In our research equivalent mutants could give a false image of the accuracy of the clusters and the overall accuracy if such a mutant was randomly selected.
PIT contains some functionality for reducing equivalent mutants\cite{pit}.
We also reduced the chance of inaccuracy by repeating our experiment.
